---
title: "Public Debates"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 3
#editor_options: 
  #chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages and data


```{r echo = TRUE, message=FALSE, warning=FALSE}
# data preprocessing
library(tidyverse)
library(tidytext)
library(lubridate)
library(tm)
library(SnowballC)
library(wordcloud)
library(qdapDictionaries)
library(reshape2) 

# data exploration
library(summarytools) # for user-friendly html summaries of data
library(ggmap) # for plotting data on a map
library(hrbrthemes)
library(showtext)
library(usmap)

showtext_auto()

# directly from google fonts
sysfonts::font_add_google("Roboto Condensed")

# set some global options
options(dplyr.width = Inf)
theme_set(theme_ipsum_rc())
```

```{r}
# Load speeches
speeches <-
  list.files(path = "data/processed/",
             pattern = "*.csv",
             full.names = TRUE) %>%
  readr::read_csv(
    id = "file_name",
    col_types = cols(
      speech_id = col_character(),
      speakerid = col_character(),
      district = col_character()
    )
  ) %>% mutate(
    chamber = factor(chamber),
    gender = factor(gender),
    party = factor(party),
    nonvoting = factor(nonvoting),
    session = as.integer(gsub(".*?([0-9]+).*", "\\1", file_name)),
    file_name = NULL
  ) %>% mutate(party = recode(party, D = 'Democratic', R = 'Republican'), 
                              name = paste(firstname, lastname)
  ) %>% filter(!is.na(speech))

```

```{r}
# Remove duplicates
speeches <- speeches %>% group_by(speech) %>% filter(n() == 1) %>% ungroup()
```

```{r}
# Add word counts for each speech
speeches <- mutate(speeches, word_count = str_count(speech ,"\\W+"))
```

```{r}
# Display statistics (not visible in browser)
speeches %>% dfSummary %>% view()
```

```{r}
speeches
```

# Speech Analysis

## Speech length frequency

```{r}
party.colors <- c(Democratic = "steelblue1", Republican = "indianred1")

# Speech length frequency
speeches  %>%
  ggplot(aes(x = word_count)) +
  geom_histogram(
    binwidth = 25,
    fill = "#69b3a2",
    color = "black",
    alpha = 0.9
  ) +
  ggtitle("Speech length frequency") +
  xlab("Speech length") +
  ylab("Count") +
  xlim(0, 1000)
```

```{r}
# Speech length frequency split by party
speeches %>%
  ggplot(aes(x = word_count, fill = party)) +
  geom_histogram(
    color = "black",
    alpha = 0.8,
    position = 'identity',
    binwidth = 25
  ) +
  xlim(0, 1000) +
  xlab("Speech length") +
  ylab("Count") +
  ggtitle("Speech length frequency by party") +
  scale_fill_manual(values=party.colors)

```
## Average speech length

```{r}
# Average speech length in session
speeches %>% group_by(session) %>% summarise(mean_words = mean(word_count)) %>%
  ggplot(aes(x = session, y = mean_words)) +
  geom_line(color = "grey") +
  geom_point(
    shape = 21,
    color = "black",
    fill = "#69b3a2",
    size = 3
  ) +
  xlab("Session") +
  ylab("Average length") +
  ggtitle("Average speech length in session")
```

```{r}
# Average speech length in session split by party
speeches %>% group_by(session, party) %>%
  summarise(mean_words = mean(word_count)) %>%
  ggplot(aes(
    x = session,
    y = mean_words,
    group = party,
    color = party
  )) +
  geom_line() +
  geom_point(
    shape = 21,
    size = 3,
    aes(fill = party),
    color = "black",
  ) + 
  xlab("Session") +
  ylab("Average length") +
  ggtitle("Average speech length in session by party")  +
  scale_fill_manual(values=party.colors) +
  scale_color_manual(values=party.colors)
```
## Total speech length
```{r}
# Total words in session
speeches %>% group_by(session) %>% summarise(sum_words = sum(word_count)) %>%
  ggplot(aes(x = session, y = sum_words)) +
  geom_line(color = "grey") +
  geom_point(
    shape = 21,
    color = "black",
    fill = "#69b3a2",
    size = 3
  ) +
  xlab("Session") +
  ylab("Total words") +
  ggtitle("Total words in session")
```

```{r}
# Total words in session split by party
speeches   %>% group_by(session, party) %>%
  summarise(sum_words = sum(word_count)) %>%
  ggplot(aes(
    x = session,
    y = sum_words,
    group = party,
    color = party
  )) +
  geom_line() +
  geom_point(
    shape = 21,
    size = 3,
    aes(fill = party),
    color = "black",
  ) +  xlab("Session") +
  ylab("Total words") +
  ggtitle("Total words in session by party") +
  scale_fill_manual(values=party.colors) +
  scale_color_manual(values=party.colors)
```

## Number of speeches

```{r}
# Number of speeches in session
speeches %>% group_by(session) %>% summarise(count = n())  %>%
  ggplot(aes(x = session, y = count)) +
  geom_line(color = "grey") +
  geom_point(
    shape = 21,
    color = "black",
    fill = "#69b3a2",
    size = 3
  ) +
  xlab("Session") +
  ylab("Speeches") +
  ggtitle("Speeches in session")
```

```{r}
# Number of speeches in session
speeches  %>% group_by(session, party) %>% summarise(count = n())  %>% ggplot(aes(
  x = session,
  y = count,
  group = party,
  color = party
)) +
  geom_line() +
  geom_point(
    shape = 21,
    size = 3,
    aes(fill = party),
    color = "black",
  ) +
  xlab("Session") +
  ylab("Speeches") +
  ggtitle("Speeches in session by party") +
  scale_fill_manual(values=party.colors) +
  scale_color_manual(values=party.colors)
```

# Speaker Analysis

## Top speakers by number of speeches

```{r}
# Top 10 speakers by number of speeches
speeches %>% group_by(speakerid) %>%
  summarise(speech_count = n(),
            party = party,
            name = name) %>%
  slice(1) %>% ungroup() %>%
  arrange(speech_count) %>% tail(10)%>% 
  mutate(name = factor(name, levels = name)) %>%
  ggplot(aes(
    x = name,
    y = speech_count,
    group = party,
    fill = party
  )) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +
  coord_flip() +
  xlab("Speaker") +
  ylab("Speeches") +
  ggtitle("Top speakers by number of speeches")  +
  scale_fill_manual(values=party.colors)
```

## Top speakers by average speech length

```{r}
# Top 10 speakers by average speech length
speeches %>% group_by(speakerid) %>%
  summarise(mean_speech_length = mean(word_count),
            party = party,
            name = name) %>%
  slice(1) %>% ungroup() %>%
  arrange(mean_speech_length) %>% tail(10)%>% 
  mutate(name = factor(name, levels = name)) %>%
  ggplot(aes(
    x = name,
    y = mean_speech_length,
    group = party,
    fill = party
  )) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +
  coord_flip() +
  xlab("Speaker") +
  ylab("Speech length") +
  ggtitle("Top speakers by average speech length") +
  scale_fill_manual(values=party.colors)
```

## Frequency of session participation

```{r}
# Frequency of session participation by Republican party
speeches %>%
  filter(party == 'Republican') %>%
  group_by(speakerid) %>%
  summarise(session_count = n_distinct(session),
            name = name) %>%
  slice(1) %>% ungroup() %>%
  ggplot(aes(x = session_count)) +
  geom_histogram(
    color = "black",
    alpha = 0.7,
    position = 'identity',
    binwidth = 1,
    fill = "indianred1"
  ) +
  xlim(0, 10) +
  xlab("Sessions") +
  ylab("Count") +
  ggtitle("Frequency of session participation by Republican party")
```

```{r}
# Frequency of session participation by Democratic party
speeches %>%
  filter(party == 'Democratic') %>%
  group_by(speakerid) %>%
  summarise(session_count = n_distinct(session),
            name = name) %>%
  slice(1) %>% ungroup() %>%
  ggplot(aes(x = session_count)) +
  geom_histogram(
    color = "black",
    alpha = 0.7,
    position = 'identity',
    binwidth = 1,
    fill = "steelblue1"
  ) +
  xlim(0, 10) +
  xlab("Sessions") +
  ylab("Count") +
  ggtitle("Frequency of session participation by Democratic party")
```

## Party switches

```{r}
speeches %>% group_by(speakerid) %>%
  mutate(party_count = n_distinct(party)) %>% filter(party_count == 2) %>% 
  ungroup %>% group_by(speakerid, session) %>% slice(1) %>% 
  select(name, party) %>% 
  ggplot(aes(
    x = session,
    y = party,
    group = name,
    color = name
  )) +
  geom_line() +
  geom_point(
    shape = 21,
    size = 3,
    aes(fill = name),
    color = "black",
  ) +
  xlab("Session") +
  ylab("Party") +
  ggtitle("Party switches")
```
```{r fig.height = 5, fig.width = 10}
speeches %>% group_by(speakerid) %>%
  mutate(party_count = n_distinct(party)) %>% filter(party_count == 2) %>% 
  ungroup %>% 
  select(name, party, date) %>% 
  ggplot(aes(date, party)) +
  geom_line() +
  geom_point(
    shape = 21,
    size = 3,
    aes(fill = party),
    color = "black",
  ) +
  facet_wrap( ~ name, ncol=1) +
  labs( NULL) +
  ggtitle("Dates of switch") +
  scale_fill_manual(values=party.colors) +
  theme(legend.position="none")
```

## US States by number of speakers

```{r}
speakers_by_states <-
  speeches %>% group_by(state) %>% 
  summarise(speaker_count = n_distinct(speakerid)) %>% ungroup()
plot_usmap(
  data = speakers_by_states,
  values = "speaker_count",
  color = "black",
  regions = c("states")
) +
  scale_fill_continuous(
    low = "white",
    high = "orange",
    name = "Number of speakers",
    label = scales::comma
  ) + theme(legend.position = "right") + 
  labs(title = "US States by number of speakers")  
```

# Word Analysis

```{r}
# Split speeches to words
speech_words <- speeches %>% unnest_tokens(word, speech)
# Remove stop words
data(stop_words)
speech_words <- speech_words %>% anti_join(stop_words)

# Filter existing words
is.word  <- function(x)
  x %in% GradyAugmented

speech_words <- speech_words[which(is.word(speech_words$word)), ]
```

## Word frequency

```{r}
# Top 10 frequent words
speech_words %>%
  count(word, sort = TRUE) %>%
  head(10) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = n, y = word)) +
  geom_col(width = 0.7,
           fill = "#69b3a2",
           color = "black",) +
  ggtitle("Word frequency") +
  xlab("Count") +
  ylab("Word")
```

```{r}
# Top 10 frequent words by Republican party
speech_words %>%
  filter(party == 'Republican') %>%
  count(word, sort = TRUE) %>%
  head(10) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = n, y = word)) +
  geom_col(
    width = 0.7,
    fill = "indianred1",
    color = "black",
    alpha = 0.7
  ) +
  ggtitle("Word frequency by Republican party") +
  xlab("Count") +
  ylab("Word")
```

```{r}
# Top 10 frequent words by Democratic party
speech_words %>%
  filter(party == 'Democratic') %>%
  count(word, sort = TRUE) %>%
  head(10) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = n, y = word)) +
  geom_col(
    width = 0.7,
    fill = "steelblue1",
    color = "black",
    alpha = 0.7
  ) +
  ggtitle("Word frequency by Democratic party") +
  xlab("Count") +
  ylab("Word")
```

```{r}
# Word Cloud
speech_words %>%
  count(word, sort = TRUE) %>%
  with(
    wordcloud(
      word,
      n,
      min.freq = 5,
      max.words = 100,
      random.order = FALSE,
      scale = c(4, .2),
      rot.per = 0.40,
      colors = brewer.pal(8, "Dark2")
    )
  )
```

```{r}
# Word cloud for Democratic party
speech_words %>%
  filter(party == 'Democratic') %>%
  count(word, sort = TRUE) %>%
  with(
    wordcloud(
      word,
      n,
      min.freq = 5,
      scale = c(4, .2),
      max.words = 100,
      random.order = FALSE,
      random.color = FALSE,
      colors = c("steelblue1", "steelblue2", "steelblue3", "steelblue")
    )
  )
```

```{r}
# Word cloud for Republican party
speech_words %>%
  filter(party == 'Republican') %>%
  count(word, sort = TRUE) %>%
  with(
    wordcloud(
      word,
      n,
      min.freq = 5,
      scale = c(4, .2),
      max.words = 100,
      random.order = FALSE,
      random.color = FALSE,
      colors = c("indianred1", "indianred2", "indianred3", "indianred")
    )
  )
```
```{r}
# Word cloud comparison by party
speech_words %>%
  count(word, party, sort = TRUE) %>%
  acast(word ~ party, value.var = "n", fill = 0) %>%
  comparison.cloud(
    scale = c(3, .2),
    colors = c("steelblue1", "indianred1"),
    max.words = 50,
    title.bg.colors = c("steelblue1", "indianred1"),
    title.colors = "white",
    title.size = 2
  )
```

## Party tf-idf

```{r}
speech_words %>% count(party, word, sort = TRUE) %>%
  ungroup() %>%
  bind_tf_idf(word, party, n) %>%
  arrange(desc(tf_idf)) %>% group_by(party) %>% slice_max(tf_idf, n = 10) %>%
  ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(tf_idf, word, fill = party)) +
  geom_col(width = 0.7,
           color = "black",
           alpha = 0.7) +
  facet_wrap( ~ party, scales = "free") +
  labs(x = "tf-idf", y = NULL) +
  ggtitle("Party tf-idf")  +
  scale_fill_manual(values = party.colors)

```

## Session tf-idf

```{r fig.height = 10, fig.width = 15}
# Session tf-idf by Republican party
speech_words %>% filter(party == 'Republican') %>% 
  count(session, word, sort = TRUE) %>%
  ungroup() %>%
  bind_tf_idf(word, session, n) %>%
  arrange(desc(tf_idf)) %>% group_by(session) %>% 
  slice_max(tf_idf, n = 5) %>%
  ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(tf_idf, word)) +
  geom_col(width = 0.7,
           color = "black",
           alpha = 0.7,
           fill = "indianred1") +
  facet_wrap( ~ session, scales = "free") +
  labs(NULL) +
  ggtitle("Session tf-idf by Republican party")

```

```{r fig.height = 10, fig.width = 15}
# Session tf-idf by Democratic party
speech_words %>% filter(party == 'Democratic') %>% 
  count(session, word, sort = TRUE) %>%
  ungroup() %>%
  bind_tf_idf(word, session, n) %>%
  arrange(desc(tf_idf)) %>% group_by(session) %>% 
  slice_max(tf_idf, n = 5) %>%
  ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(tf_idf, word)) +
  geom_col(width = 0.7,
           color = "black",
           alpha = 0.7,
           fill = "steelblue1") +
  facet_wrap( ~ session, scales = "free") +
  labs(NULL) +
  ggtitle("Session tf-idf by Democratic party")

```
## Sentiment analysis

```{r}
# Word cloud comparison
speech_words %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(
    colors = c("#F8766D", "#00BFC4"),
    max.words = 50,
    title.bg.colors = "white"
  )
```

```{r}
# Word contributions
speech_words %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(word) %>%
  summarize(occurences = n(),
            contribution = sum(value))  %>%  slice_max(abs(contribution), n = 20) %>%
  mutate(word = reorder(word, contribution)) %>%
  ggplot(aes(contribution, word, fill = contribution > 0)) +
  geom_col(
    width = 0.7,
    color = "black",
    alpha = 0.7,
    show.legend = FALSE
  ) +
  labs(y = NULL) +
  ggtitle("Word contribution") +
  xlab("Contribution")
```

```{r}
# Word contributions by party
speech_words %>%
  count(party, word, sort = TRUE) %>%
  ungroup() %>%
  inner_join(get_sentiments("afinn")) %>%
  mutate(contribution = value * n / sum(n)) %>%
  group_by(party) %>%
  slice_max(abs(contribution), n = 20) %>%
  ungroup() %>%
  mutate(word = reorder_within(word, contribution, party)) %>%
  ggplot(aes(contribution, word, fill = contribution > 0)) +
  geom_col(
    width = 0.7,
    color = "black",
    alpha = 0.7,
    show.legend = FALSE
  ) +
  facet_wrap( ~ party, scales = "free") +
  scale_y_reordered() +
  labs(x = "Contribution",
       y = NULL) +
  ggtitle("Word contribution by party")

```

